An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-03-29 EconGrowthSeminar.Rnw
Now including Caleb's inputs.

\begin{frame}{Economic (Growth) Modeling} %----------------------------
  \begin{center}
    \includegraphics[height=2.5cm]{figure_other/Growth_Models.pdf}
  \end{center}
  \begin{itemize}
    \item{Goal: predict economic output over time}
    \item{Economic output ($Y$) is a function of ``factors of production''}
    \item{Output ($Y$) measured by GDP}
    \item{All inputs and outputs normalized to initial year: $y = Y/Y_0, e = E/E_0$, etc.}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Economic Data: Definitions} %----------------------------
  \begin{itemize}
    \item{Output ($Y$): the real market value of the total amount of produced goods in an economy (GDP).}
    \item{Capital Stock ($K$): the real market value of the total amount of goods in an economy used in the production process of output (machines, tools, infrastructure, etc.).}
    \item{Labor ($L$): the number of workers in an economy that produce output. Labor can be treated as either the total number of workers in an economy or the total number of hours worked per year in an economy.}
  \end{itemize}
\end{frame}

\begin{frame}{Sources of Economic Data} %----------------------------
  \begin{center}
    \scriptsize
    \begin{tabular}{r c|c l}
      \multirow{3}{*}{WB GEP} & \multirow{3}{*}{\includegraphics[scale=0.07]{figure_other/Figure_World_Bank.png}} & $Y$, $K$ for all countries & \multirow{3}{*}{Real 2005\$} \\
                              &                          & (1980-2011: US, UK, JP; & \\
                              &                          & 1991-2011: CN, ZA, SA, IR, TZ, ZM) & \\
      \hline
      \multirow{4}{*}{WB WDI} & \multirow{4}{*}{\includegraphics[scale=0.07]{figure_other/Figure_World_Bank.png}} & Pop, Pop < 15 yo,    & \multirow{4}{*}{Total Employment} \\
                              &                          & Emp/Pop ratio > 15 yo, &                                \\
                              &                          & used to find $L$, (1991-2011: &                         \\
                              &                          & CN, ZA, SA, IR, TZ, ZM) &                         \\
      \hline
      \multirow{2}{*}{US BLS} & \multirow{2}{*}{\includegraphics[scale=0.07]{figure_other/Figure_BLS.png}} & $L$ for developed countries & Labor hours \\
                              &                          & (1980-2011: US, UK, JP)     & per year \\ 
    \end{tabular}
  \normalsize
  \end{center}
\end{frame}

@@

\begin{frame}{Energy Data: Definitions} %----------------------------
  \begin{itemize}
    \item{Thermal Energy ($Q$): The heat content of an energy carrier.}
    \item{Exergy ($X$): The maximum work that can be extracted from a heat carrier by a process that brings the system into equillibrium with a thermal reservoir (typically the surrounding environment).}
    \item{Useful Work ($U$): Energy that directly generates economic activity.}
  \end{itemize}
\end{frame}

\begin{frame}{Energy Flows} %----------------------------
  Caleb's diagram here
\end{frame}

\begin{frame}{Sources of Energy Data} %----------------------------
  Table here.
\end{frame}

@@

\begin{frame}{$SSE$ Minimization}{US single-factor with $Q$} %---------------------------- 
<<3D_Graph_US_SF_Q, fig.width=presentationGraph1ColWidth*1.3, fig.height=presentationGraph1RowHeight*1.3, eval=TRUE>>=
sf3DSSEGraph("US", "Q")
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{US}{0.055}  
\end{frame}

@@

\begin{frame}{Benefit to Including Energy?}  %----------------------------
<<US_Partial_Residuals_Plot, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight>>=
partialResidualPlot("US", "Q", textScaling=textScaleForGraphs)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{US}{0.055}
\end{frame}

\begin{frame}{Benefit to Including Energy?}  %----------------------------
<<UK_Partial_Residuals_Plot, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight>>=
partialResidualPlot("UK", "Q", textScaling=textScaleForGraphs)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{UK}{0.055}
\end{frame}

>>>>

\begin{frame}{Economic (Growth) Modeling} %----------------------------
  \begin{center}
    \includegraphics[height=2.5cm]{figure_other/Growth_Models.pdf}
  \end{center}
  \begin{itemize}
    \item{Goal: predict economic output over time}
    \item{Economic output ($Y$) is a function of ``factors of production''}
    \item{Output ($Y$) measured by GDP}
    \item{All factors and outputs indexed to initial year: $y = Y/Y_0, e = E/E_0$, etc.}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Economic Definitions} %----------------------------
  \begin{itemize}
    \item{Output ($Y$): the real market value of the total amount of produced goods in an economy (GDP).}
    \item{Capital Stock ($K$): the real market value of the total amount of goods in an economy used in the production process of output (machines, tools, infrastructure, etc.).}
    \item{Labor ($L$): the number of workers in an economy that produce output. Labor can be treated as either the total number of workers in an economy or the total number of hours worked per year in an economy.}
  \end{itemize}
\end{frame}

\begin{frame}{Sources of Economic Data} %----------------------------
  \begin{center}
    \scriptsize
    \begin{tabular}{r c c l}
      \multirow{3}{*}{WB GEP} & \multirow{3}{*}{\includegraphics[scale=0.07]{figure_other/Figure_World_Bank.png}} & $Y$, $K$ for all countries & \multirow{3}{*}{Real 2005\$} \\
                              &                          & (1980-2011: Developed; & \\
                              &                          & 1991-2011: Other) & \\
      \hline
      \multirow{4}{*}{WB WDI} & \multirow{4}{*}{\includegraphics[scale=0.07]{figure_other/Figure_World_Bank.png}} & Pop, Pop < 15 yo,    & \multirow{4}{*}{Total Employment} \\
                              &                          & Emp/Pop ratio > 15 yo, &                                \\
                              &                          & used to find $L$, (1991-2011: &                         \\
                              &                          & CN, ZA, SA, IR, TZ, ZM) &                         \\
      \hline
      \multirow{2}{*}{US BLS} & \multirow{2}{*}{\includegraphics[scale=0.07]{figure_other/Figure_BLS.png}} & $L$ for developed countries & Labor hours \\
                              &                          & (1980-2011: US, UK, JP)     & per year \\ 
    \end{tabular}
    \normalsize
  \end{center}
\end{frame}

@@

\begin{frame}{Energy Definitions} %----------------------------
  \begin{itemize}
    \item{Thermal Energy ($Q$): The heat content of an energy carrier.}
    \item{Exergy ($X$): The maximum work that can be extracted from a heat carrier by a process that brings the system into equillibrium with a thermal reservoir (typically the surrounding environment).}
    \item{Useful Work ($U$): Energy that directly generates economic activity.}
  \end{itemize}
  \begin{center}Note: indexed $Q$ and $X$ are identical unless energy mix changes significantly. \end{center}
\end{frame}

\begin{frame}{Energy Flows} %----------------------------
  \begin{center}
    \includegraphics[scale=0.4]{figure_other/Figure_Energy_Sankey.pdf}
  \end{center}
\end{frame}

\begin{frame}{Sources of Energy Data} %----------------------------
  \begin{center}
    \scriptsize
    \begin{tabular}{r c c l}
      \multirow{3}{*}{IEA} & \multirow{3}{*}{\includegraphics[scale=0.20]{figure_other/Figure_IEA.png}} & Coal, Oil, Petroleum, & \multirow{3}{*}{GJ/year} \\
                              &                          & Nat. gas, Nuclear, & \\
                              &                          & Renewables & \\
      \hline
      \multirow{4}{*}{FAOStat} & \multirow{4}{*}{\includegraphics[scale=0.07]{figure_other/Figure_FAOStat.png}} & Fuel wood,    & \multirow{4}{*}{GJ/year} \\
                              &                          & Available livestock &                                \\
                              &                          & (1980-2011: Developed; &                         \\
                              &                          & 1991-2011: Other) &                         \\
      \hline
      \multirow{2}{*}{REXS} & \multirow{2}{*}{\includegraphics[scale=0.05]{figure_other/Figure_Warr.png}} & Energy allocation factors. &  \\
                              &                          & (1980-2000: Developed)     & \\ 
    \end{tabular}
    \normalsize
  \end{center}
\end{frame}

@@

\begin{frame}{$SSE$ Minimization}{US single-factor with $Q$} %---------------------------- 
<<3D_Graph_US_SF_Q, fig.width=presentationGraph1ColWidth*1.3, fig.height=presentationGraph1RowHeight*1.3>>=
sf3DSSEGraph("US", "Q")
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{US}{0.055}  
\end{frame}

@@

\begin{frame}{Beneficial to Include Energy?}  %----------------------------
<<US_Partial_Residuals_Plot, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight>>=
partialResidualPlot("US", "Q", textScaling=textScaleForGraphs)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{US}{0.055}
\end{frame}

\begin{frame}{Beneficial to Include Energy?}  %----------------------------
<<UK_Partial_Residuals_Plot, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight>>=
partialResidualPlot("UK", "Q", textScaling=textScaleForGraphs)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{UK}{0.055}
\end{frame}
